Dark matter halo shape in the Auriga simulations: radial dependence, time evolution and correlation with baryon disk properties

### 2019-07-15 Document/paper.tex
new plot on the cumulative alignment

\title[The shape of dark matter haloes in the Auriga simulations]
{Dark matter halo shape in the Auriga simulations: radial
  dependence, time evolution and correlation with baryon disk
  properties}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg
Germany.\\
}

@@

%\begin{figure*}
%\centering
%{\includegraphics[width=0.6\textwidth]{./pics/MHD_Vs_DM/halo16_new.png}}
%\caption{Radial profile for the axial ratios $b/a$, $c/a$, and the triaxiality parameter
%  $T$ at redshift $z=0$.
%  The continous/dashed lines correspond to the MHD/DM simulations respectively.
%  The vertical line corresponds to the virial radius.}
%\label{fig:triaxial_radius}
%\end{figure*}

@@

@@

The DM halo shape at a fixed radius is an estimate of either
the isopotential or isodensity surfaces.  
Observational inference models usually estimate the 
isopotential contours which are probed by tracers (gas, stars), while
simulations work with the isodensity contours which can be directly
calculated from particle positions.  
Furthermore, the density contours in thin shells are very sensitive to
the presence of small satelites.  
For this reason we measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.  
This method yields results in good agreement to the isodensity
contours for radii smaller than half the virial radius as explored by
\citep{Vera-Ciro_et_al._2011}.

@@

\begin{figure*}
  \includegraphics[width=\columnwidth]{./pics/Redshift/halo_27_DM_Z_correlation.png}
  \includegraphics[width=\columnwidth]{./pics/Redshift/halo_24_level3_MHD_Z_Triax.png}
  \caption{Radial profile at $z=0$ (dashed line) and 
    the historical profile at the virial radius (colour line). 
    Results for DM simulations on the left, for MHD on the right.{\bf
      Update with halo 24}}
  \label{fig:redshift_triaxial}
\end{figure*}

\begin{figure*}
  \subfloat[halo 16 DM]{\includegraphics[width=\columnwidth]{./pics/Redshift/halo_16_level3_DM_Z.png}}
  \subfloat[halo 16 MHD]{\includegraphics[width=\columnwidth]{./pics/Redshift/halo_16_level3_MHD_Z.png}}
  \caption{Radial profile (comoving) of axial ratios for halo 16 in
    terms of redshift (color). This halo maintains its shape until
    $z\approx 1$ obviating the systematic rounding effect in time from
    asymmetric potentials. Each plot-line represents the radial
    profile at a determined redshift. Conclussion: Mainly for b/a, it increases for:
    (1) for bigger radii (fixed redshift)(2) for lower redshift (fixed
    radius) It explains the correlation between radial profile and
    history, but does not require that curves match in the triaxial
    plane. {\bf En esta grafica el panel de abajo deberia ser la triaxialidad.}}
  \label{fig:RedshiftGood}
\end{figure*}

We know that DM-only have a steady and monotonous
trend towards larger sphericity with increasing radius. 
This radial trend is mimicked when the shape is measured at the virial
radius as a function of cosmic time.
The halo should become rounder with decreasing redshift, this is
expected by the continuous influence of the gravitational potential on
a collisionless fluid \citep{Vera-Ciro_et_al._2011}.

We show in the left panel of Figure \ref{fig:redshift_triaxial} an
example with this effect in Halo 27 of the DM simulations:
the radial triaxiality at $z=0$ and the historical triaxiality at
$R_{\rm vir}$ are correlated. 
The right panel on the same Figure \ref{fig:redshift_triaxial}
illustrates how this correlation is absent in MHD simulations.

This means that for DM-only halos one can approximate its shape at
higher redshift by simply sampling its 
current shape at a smaller radius. 
This correlation seems to be prompted by the continous inside-out
build-up of the dark matter halo; baryonic effects seem to jeopardize
the apparent smoother growth in DM-only simulations.

%Discuss if this correlation may be recovered if compared for example at Disk radius in stead of virial radius.\\

>>>>

\title[The shape of dark matter haloes in the Auriga simulations]
{Dark matter halo shape in the Auriga simulations: radial
  dependence, time evolution and correlation with baryon disk
  properties}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg
paperGermany.\\
}

@@

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{cumulative_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{cumulative_alignment_MHD.pdf}
\end{center}
\caption{Cumulative alignment}
\label{fig:cumulative_alignment}
\end{figure*}

\begin{figure*}
\begin{center}
  \subfloat[DM shape at small radius]{\includegraphics[width=0.4\textwidth]{Z_Triax_level3_set_A_DM.pdf}} 
  \subfloat[DM shape at large radius]{\includegraphics[width=0.4\textwidth]{Z_Triax_level3_set_A_MHD.pdf}} 
\end{center}
\caption{}
\label{fig:triaxial_history}
\end{figure*}

\begin{figure*}
\begin{center}
  \subfloat[DM shape at small radius]{\includegraphics[width=0.4\textwidth]{Z_Triax_level3_set_B_DM.pdf}} 
  \subfloat[DM shape at large radius]{\includegraphics[width=0.4\textwidth]{Z_Triax_level3_set_B_MHD.pdf}} 
\end{center}
\caption{}
\label{fig:triaxial_history}
\end{figure*}

@@

The DM halo shape at a fixed radius is an estimate of either
the isopotential or isodensity surfaces.  
Observational inference models usually estimate the 
isopotential contours which are probed by tracers (gas, stars), while
simulations work with the isodensity contours which can be directly
calculated from particle positions.  
Furthermore, the density contours in thin shells are very sensitive to
the presence of small satelites.  
For this reason we measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.  
This method yields results in good agreement to the isodensity
contours in the inner halo region, roughly for radii smaller than one
quarter the virial radius as explored by
\citep{Vera-Ciro_et_al._2011}.

@@

In the previous section we stablished that DMO halos at $z=0$ have
a steady and monotenouse trend towards decreasing triaxility with
increasing radius.
\cite{Vera-Ciro_et_al._2011} found that this  radial trend at $z=0$ 
correlates with the halo shape as a function of cosmic time, i.e.
the halo becomes less triaxial as cosmic time progresses. 
This correlation was interpreted as each radial shell retaining memory
of the accreting conditions at the virial radius at the time of
collapse.

Here we use the six \texttt{level3} halos to measure to what extent
the presence of baryons affect this memory effect.
The left column of Figure \ref{fig:redshift_triaxial} shows the six
halos in the triaxility plane. 
The colour dots show the time evolution while the black line
corresponds to the radial trend at $z=0$.
The close overlap of the spatial and temporal trends confirms the
memory effect found by \cite{Vera-Ciro_et_al.2011}.

In the MHD sample the memory effect is less evident. 
The right column of Figure \ref{fig:redshif_triaxial} shows the
corresponding results for the six halos in the MHD sample.

%Discuss if this correlation may be recovered if compared for example
%at Disk radius in stead of virial radius.\\
